\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-08-14 manuscript.Rnw
Typo fixes

Identifiers for the complete array of transitions matrices (inclusive of SOC and the new interventions) are generated with \code{tpmatrix_id()}.

@@

A design decision that is worth commenting on is the use of \pkg{R6}. While it may not be ``traditional'' \proglang{R}, we believe it fits our modular design nicely. It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that it is natural to have classes representing separate submodels for disease progression, costs, and utility, which can be combined into a single class representing an economic model. A particularly important benefit of \pkg{R6} is that methods can belong to objects. In our framework, this means that submodels can not only contain inputs for the economic model, but can also generate outputs. For example, the disease model of a PSM not only contains the parameters of the survival models and the input data required for prediction, but can also generate relevant outcomes such as survival curves or hazard functions. Likewise, the disease model of an iCTSTM not only stores parameters and input data, but can be used to simulate trajectories through a multi-state model and compute state probabilities. This makes it easier to check that the submodels produce desired outputs, compare submodels with different underlying modeling assumptions (e.g., different parametric survival distributions), and to use the package in a more modular way. The latter may be especially relevant to users who only want to use some features of the package; for, instance a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels. All that said, we utilize \code{S3} whenever it makes sense as it is more familiar to most \proglang{R} users.

@@

\pkg{mstate} is another package for multi-state modeling that is designed for estimation of non-parametric and semi-parametric (i.e., Cox proportional hazards) models \citep{de2011mstate}. It can predict state probabilities and simulate multi-state trajectories (to facilitate prediction for clock-reset models) from cumulative hazard functions evaluated at specific times. \citet{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications. There are however, some disadvantages of using \pkg{mstate} for health economic modeling since it was not designed for it. \code{mssample()}---the core function for simulating multi-state models---does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output. It is also less computationally efficient than \pkg{hesim} for two main reasons. First, simulation of general cumulative hazards functions is considerably slower than simulation of parametric distributions when efficient random number generators are available. Second, it is written completely in \proglang{R} and incorporating PSA, multiple treatment strategies, or heterogeneous patients results in code that is not vectorized. To compare computational performance (see ``Benchmarks'' on the package website), we ran 1,000 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration. At the time of writing, the model completed in 2.8 seconds in \pkg{hesim} and 3.1 hours in \pkg{mstate}, showing that computational considerations are not just an academic concern.

@@

There are of course features in other \proglang{R} packages that are not in the current version of \pkg{hesim}. For example, \pkg{heemod} has an explicit function, \code{calibrate_model()}, for calibrating model parameters that are difficult to measure. Furthermore, tunnel states can be automatically created (rather than needing to be created manually) with \pkg{heemod} using the \code{state_time} variable so that costs or transition probabilities can very as a function of time in a health state. \pkg{hesim} is also limited in its support for value of information analyses. In particular, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI}, and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}. In this respect, these packages are complementary and simulation output from \pkg{hesim} could be used as inputs to their more advanced value of information functions.

>>>>

Identifiers for the complete array of transition matrices (inclusive of SOC and the new interventions) are generated with \code{tpmatrix_id()}.

@@

A design decision that is worth commenting on is the use of \pkg{R6}. While it may not be ``traditional'' \proglang{R}, we believe it fits our modular design nicely. It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that it is natural to have classes representing separate submodels for disease progression, costs, and utility, which can be combined into a single class representing an economic model. A particularly important benefit of \pkg{R6} is that methods can belong to objects. In our framework, this means that submodels can not only contain inputs for the economic model, but can also generate outputs. For example, the disease model of a PSM not only contains the parameters of the survival models and the input data required for prediction, but can also generate relevant outcomes such as survival curves or hazard functions. Likewise, the disease model of an iCTSTM not only stores parameters and input data, but can be used to simulate trajectories through a multi-state model and compute state probabilities. This makes it easier to check that the submodels produce desired outputs, compare submodels with different underlying modeling assumptions (e.g., different parametric survival distributions), and to use the package in a more modular way. The latter may be especially relevant to users who only want to use some features of the package; for instance, a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels. All that said, we utilize \code{S3} whenever it makes sense as it is more familiar to most \proglang{R} users.

@@

\pkg{mstate} is another package for multi-state modeling that is designed for estimation of non-parametric and semi-parametric (i.e., Cox proportional hazards) models \citep{de2011mstate}. It can predict state probabilities and simulate multi-state trajectories (to facilitate prediction for clock-reset models) from cumulative hazard functions evaluated at specific times. \citet{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications. There are, however, some disadvantages of using \pkg{mstate} for health economic modeling since it was not designed for it. \code{mssample()}---the core function for simulating multi-state models---does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output. It is also less computationally efficient than \pkg{hesim} for two main reasons. First, simulation of general cumulative hazards functions is considerably slower than simulation of parametric distributions when efficient random number generators are available. Second, it is written completely in \proglang{R} and incorporating PSA, multiple treatment strategies, or heterogeneous patients results in code that is not vectorized. To compare computational performance (see ``Benchmarks'' on the package website), we ran 1,000 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration. At the time of writing, the model completed in 2.8 seconds in \pkg{hesim} and 3.1 hours in \pkg{mstate}, showing that computational considerations are not just an academic concern.

@@

There are of course features in other \proglang{R} packages that are not in the current version of \pkg{hesim}. For example, \pkg{heemod} has an explicit function, \code{calibrate_model()}, for calibrating model parameters that are difficult to measure. Furthermore, tunnel states can be automatically created (rather than needing to be created manually) with \pkg{heemod} using the \code{state_time} variable so that costs or transition probabilities can vary as a function of time in a health state. \pkg{hesim} is also limited in its support for value of information analyses. In particular, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI}, and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}. In this respect, these packages are complementary and simulation output from \pkg{hesim} could be used as inputs to their more advanced value of information functions.
